---
title: "Lab8-Block Treaments"
author: "Katherine Filpo"
date: "2/28/2020"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
SAPLINGS <- read.csv("~/Documents/BioStats/lab8/SAPLINGS.csv")
#View(SAPLINGS)
```

## Question 1: Is length explained by water?

H~0~ There is no relationship between water treatment and length of plant. 

H~A~ The water treament will affect the length of plant, in the manor that the more water, the longer the plant. 

```{r}
model_1<-aov(SAPLINGS$FINALHT ~factor(SAPLINGS$WATER_TREAT))
summary(model_1)
```

#### Plot of the water treament and sapling length

```{r pressure, echo=FALSE}
Fig1 <- ggplot(SAPLINGS, aes(x=factor(SAPLINGS$WATER_TREAT), y=SAPLINGS$FINALHT))     +geom_boxplot() + xlab("Water Treatment") + ylab("Final Length (cm)")

Fig1
```

**Figure 1.** The water treatment for seems to have a weak effect on the final length of the plats but wothout taking into effect the initial length, it is not possible to conclude without a doubt that it is only due to the water treatment, so we cannot accept or reject the hypothesis. (*F*~(3,36)~ = 1.9721) (lower the f the higher the p)

## Question 2: Testing the initial height

H~0~ The initial height of the plant has no relationship on the final length of the plant. 

H~A~ THe initial height of the plant has an effect on the final length of the plant. 

```{r}
model_2<-aov((SAPLINGS$FINALHT-SAPLINGS$INITHT)~factor(SAPLINGS$WATER_TREAT))
summary(model_2)
```
## Question 3: Two way Anova

H~0~ The initial height has no effect on the final height of the plants even when taking into account the wter treatment. 

H~A~ The iniital height and the water treatment both have an effect on the final height of the plants. 

Here the initial height is a block so it must be in the second half of the data when typed; its basically saying is final height explained by water treatment when taking into account the initial height? 
```{r}
model_3<-aov(SAPLINGS$FINALHT~ SAPLINGS$INITHT + factor(SAPLINGS$WATER_TREAT))
summary(model_3)
```

#### Plot of the effect of water treament on height when taking into account the initial height (two way anova)

```{r, echo=FALSE}
Fig2 <- ggplot(model_3, aes(x=factor(SAPLINGS$WATER_TREAT), y=(SAPLINGS$FINALHT-SAPLINGS$INITHT)))     +geom_boxplot() + xlab("Water Treatment") + ylab("Final Length (cm)")

Fig2
```

**Figure 2.** The water treatment and the initial length of the plant have an effect the final length of the plant (*F*~(3,36)~ = 16637.378).

## Conclusion
Taking into account the initial legnth and the water treament, there is an effect of the 
water treatment so the H~A~ is supported. 
