---
title: "Bio244 - LAB 6 - ANOVA"
author: "Gabriel Perron"
date: "2/13/2019"
output: html_document
---

```{r loading necessary datasets and packages, echo=FALSE, results="hide"}
#Before starting, I import my data
TREE_S15 <- read.csv("~/Documents/BioStats/lab6/Tree_2015 - Sheet1.csv")
#View(TREE_S15)
TREE_F17 <- read.csv("~/Documents/BioStats/lab6/Tree_2017 - Sheet1.csv")
#TREE <- read.csv("~/Dropbox/1_University/2_Education/1_Teaching/1_Bard/1_Courses/2_Biostats/0_Data_set/TREE.csv")
library(ggplot2)

```


#Objective
Using t-test to investigate whether the average tree height in the RKC parking lots differ in heights between the different rows and between different years.  

##Question 1

See board.


##Question 2

###Hypotheses
H~0~: There is no effect of row on height in year 2015.  
(or there is not difference in the average height of tree among row A, B, and C)  
H~A~: There is an effect of row on height in year 2015. 
(or there is a difference in the average height of tree among row A, B, and C).

```{r Q2}
model_1 <- aov(TREE_S15$Height ~ TREE_S15$Row)
summary(model_1)

```
Conclusion: We reject H~0~; there is an effect of row on tree height (*F*~(2,15)~ = 4.516; *P* = 0.0292).

##Question
```{r}
Fig1 <- ggplot(data=TREE_S15, aes(x=TREE_S15$Row, y=TREE_S15$Height)) + 
        geom_boxplot()
Fig1
```
  
**Figure 1.** There is an effect of row on tree height (*F*~(2,15)~ = 4.516; *P* = 0.0292).

##Question 4

###Hypotheses
H~0~: There is no effect of row on height in year 2017.
H~A~: There is an effect of row on height in year 2017. 

```{r Q2}
model_2 <- aov(TREE_F17$Height ~ TREE_F17$Row)
summary(model_2)

```
Conclusion: We fail to reject H~0~; there is no effect of row on tree height (*F*~(2,15)~ = 1.491; *P* = 0.257).


```{r}
Fig2 <- ggplot(data=TREE_F17, aes(x=TREE_F17$Row, y=TREE_F17$Height)) + 
        geom_boxplot()
Fig2
```
  
  **Figure 2.** There is no effect of row on tree height (*F*~(2,15)~ = 1.491; *P* = 0.257).
