---
title: "Lab6-ANOVA and Variance"
author: "Katherine Filpo"
date: "2/19/2020"
output:
  html_document:
    df_print: paged
---

```{r}
TREE_S15 <- read.csv("~/Documents/BioStats/lab6/Tree_2015 - Sheet1.csv")
#View(TREE_S15)
TREE_F17 <- read.csv("~/Documents/BioStats/lab6/Tree_2017 - Sheet1.csv")
#View(TREE_F17)
library(ggplot2)
```

H~0~ My null hypothesis is that there is no differnce between the height of the trees. 

H~A~ My actual hypothesis is that there is a difference in the height of the trees. 

## Bonferroni Correction 
Since there are many factors of variation to consider, I would say that my new conclusions are that 

corrected P would be 0.05/3 which is 0.01667, as there are tree nteractions happening in this experiment, between A and B, A and C, and B and C.

With this new conclusion, I would say that my conclusion did change, as none of the t values I got would be significant, or at least I would say so without looking more closely at all the p values.

```{r}
p.adjust(0.02770,n=3,method='bonferroni')
#
```


## Q2 Analysis of Variance 

An anova of the 2015 data and the plot of it. 

```{r}
tree.aov <- aov(TREE_S15$Height ~ TREE_S15$Row)
anova(tree.aov)
summary(tree.aov) #does the same thingas anova
print(tree.aov)
```

```{r}
ggplot(TREE_S15, aes(x = Row, y = Height)) +geom_point()+geom_boxplot()
```
Fail to regict H~0~ but I dont think i get why
Figure1.

## Q3 

An anova of the data from 2017 and the plot of it.

```{r}
tree1.aov <- aov(TREE_F17$Height ~ TREE_F17$Row)
anova(tree1.aov)
```
```{r}
ggplot(TREE_F17, aes(x = Row, y = Height)) +geom_point()+geom_boxplot()
```
also fail to reject H~0~
Figure2.There us no ffect of row on tree height(F(2,15) = 1.491; P=257)

## Final TUKEY analysis

```{r}
Tukey <- TukeyHSD(x=tree.aov, 'TREE_S15$Row', conf.level = 0.95)
#parenthesis automatically print 
plot(Tukey , las=1, col="brown")
#this to help test that you are doing the tests 
#confidence intervals 
```

