---
title: "Bio244 - Practical I"
author: "Gabriel Perron"
date: "2020/4/01"
output:
  html_document: default
  pdf_document: default
---


```{r, echo=FALSE, results="hide",message=FALSE,warning=FALSE}
#loading necessary libraries for running the script
library(ggplot2)
SUNFLOWER <- read.csv("~/Dropbox/1_University/2_Education/1_Teaching/1_Bard/1_Courses/2_Biostats/0_Data_set/SUNFLOWER.csv")
```

**a) Word Equation:** BIOMASS ~ FARM + WATER + DENSITY + WATER:DENSITY

**b) Hypothesis:**

*Interaction*  
**H~0i~*: There is no effect of the interaction between WATER and DENSITY on BIOMASS;  
**H~Ai~*: There is an effect of the interaction between WATER and DENSITY on BIOMASS;  
  
*Density*  
**H~0density~*: There is no effect of DENSITY on BIOMASS;  
**H~Adensity~*: There is an effect of DENSITY on BIOMASS;  

*Water*  
**H~0water~*: There is no effect of WATER on BIOMASS;  
**H~Awater~*: There is an effect of WATER on BIOMASS;  

*Farm*  
**H~0farm~*: There is no effect of FARM on BIOMASS;  
**H~Afarm~*: There is an effect of FARM on BIOMASS;  

**c) Statistical Analysis**  
```{r}
model_seeds <- lm(SUNFLOWER$BIOMASS~factor(SUNFLOWER$FARM)+factor(SUNFLOWER$WATER)+factor(SUNFLOWER$DENSITY)+factor(SUNFLOWER$WATER):factor(SUNFLOWER$DENSITY))
anova(model_seeds)
```

**d) Conclusion:**   
We reject H~0i~. There is an effect of the interaction between WATER and DENSITY on BIOMASS (*F*~(4,16)~ = 5.77; *P* = 0.005).  
We rejext H~0farm~. There is no effect of FARM on BIOMASS (*F*~(2,16)~ = 3.88; *P* = 0.04).


**e) Data Vizualisation:**  
```{r}
(Figure1 <- ggplot(SUNFLOWER, aes(y=BIOMASS, x=factor(DENSITY), fill=factor(WATER))) +
            geom_boxplot() +
            xlab("Density treatment") +
            ylab("Biomass (g)")+
            theme_classic())
```
 
 **Figure 1.** After accounting for the effect of farm on sunflower seeds biomass (*F*~(2,16)~ = 3.88; *P* = 0.04), we found a significant effect of the interaction between WATER and DENSITY on BIOMASS (*F*~(4,16)~ = 5.77; *P* = 0.005).






