---
title: "Practical"
author: "Katherine Filpo"
date: "4/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
SUNFLOWER <- read_csv("SUNFLOWER.csv")
#View(SUNFLOWER)
```

***Word Equation***

Biomass ~ light + water treatment

***Hypotheses***

H~01~ The water treatment will have no effect on the biomass of the sunflowers.  

H~A1~ The water treatment will have an effect on the biomass of the sunflowers.

H~02~ The amount of sunlight will have no effect on the biomass of the sunflowers. 

H~A2~ The amount of sunlight will have an effect on the biomass of the sunflowers.

H~03~ The light and water treatment will not have an effect on the biomass of the sunflowers. 

H~A3~ The light and water treatment will not have an effect on the biomass of the sunflowers. 

### Analysis of Water Treatment

```{r}
model_1 <-aov(SUNFLOWER$BIOMASS ~ SUNFLOWER$WATER)
summary(model_1)
```

### Analysis of Light Treatment

```{r}
model_2 <-aov(SUNFLOWER$BIOMASS ~ SUNFLOWER$DENSITY)
summary(model_2)
```

### Analysis of Both Water and Light Treatment

```{r}
model_3 <-aov(SUNFLOWER$BIOMASS ~ (SUNFLOWER$DENSITY+SUNFLOWER$WATER))
summary(model_3)
```

## Figures

```{r, echo=FALSE}
Fig1 <- ggplot(SUNFLOWER, aes(x=factor(SUNFLOWER$WATER), y=SUNFLOWER$BIOMASS)) +
  geom_boxplot() + 
  theme_classic() +
  xlab("Water Treatment") + 
  ylab("Biomass of Plant")

Fig1
```

***Figure 1.*** The biomass of plants are higher when the water treatment is similarly greater, supporting the H~A1~, **F**(1,25)= 21.56.

```{r, echo=FALSE}
Fig2 <- ggplot(SUNFLOWER, aes(x=factor(SUNFLOWER$DENSITY), y=SUNFLOWER$BIOMASS)) +
  geom_boxplot() + 
  theme_classic() +
  xlab("Density of Plants") + 
  ylab("Biomass of Plant")

Fig2
```

***Figure 2.*** The biomass of the plants are higher when the density of the plants is lower but the relationship seems to be weak and the H~02~ must be accepted, **F**(1,25)= 4.052. 


## Conclusion
The best way to get the most biomass of the sunflower seeds is to have a higher water treatment and have less density of plants **F**(1,24)= 8.423. 

