---
title: "Bio244 - LAB 15 - Multiple Regression"
author: "Gabriel Perron"
date: "2019/4/01"
output:
  html_document: default
  pdf_document: default
---


```{r loading necessary datasets and packages, echo=FALSE, results="hide"}
#Before starting, I import my data
whale <- read.csv("~/Dropbox/1_University/2_Education/1_Teaching/1_Bard/1_Courses/2_Biostats/0_Data_set/whale.csv", header=TRUE)
library(car)
```

Before we go aheaddd with this kind of analysis, I like to use the "attach" function so that I don't have to refer to the dataset each time I try/use a variable:  
```{r}
attach(whale)
names(whale)
```

#Question 1
```{r}
model_whale_1 <- lm(LRGWHAL~YEAR*MONTH*DAY*CLOUD8AM*VIS8AM)
summary.aov(model_whale_1)
```
CLOUD8AM and VIS8AM are significant.  
  
#Question 2
```{r}
model_whale_2 <- lm(LRGWHAL~YEAR+MONTH+DAY+CLOUD8AM+VIS8AM)
summary.aov(model_whale_2)
```
Again, CLOUD8AM and VIS8AM are significant.  
  

#Question 3
```{r}
model_whale_3 <- lm(LRGWHAL~VIS8AM+CLOUD8AM+DAY+MONTH+YEAR)
summary.aov(model_whale_3)
```
This time, only VIS8AM is significant!
  
  
#Question 4
Running the same models, this time using the "Anova" function with type II sums of squares. This type of sums of squares, basically assumes that each factor is always last.
```{r}
Anova(lm(LRGWHAL~YEAR+MONTH+DAY+CLOUD8AM+VIS8AM), type=2)
```
VIS8AM is sifgnificant.

#Question 5
```{r}
Anova(lm(LRGWHAL~VIS8AM+CLOUD8AM+DAY+MONTH+YEAR),type=2)
```
Again, VIS8AM is significant. In fact, VIS8AM has the same F-value and P-value irregardless of its position in the mode.


#Question6
```{r}
(box.whale <- ggplot(whale, aes(factor(YEAR),LRGWHAL))+
                    geom_boxplot(outlier.size=1,lwd=0.5,fatten = 1.25)  +
                    geom_point(aes(fill = YEAR), size = 3, shape = 21, position = position_jitterdodge()) +
                    theme_classic() )
```

#Question 7
No. Looking at the interaction plot above, it is clear the number of whales does not follow a linear trend over time (e.g., it is not increasing or decreasing). Year should be considered as a categorical variable rather.

#Question 8
```{r}
Anova(lm(LRGWHAL~factor(YEAR)+MONTH+DAY+CLOUD8AM+VIS8AM), type=2)
```
VIS8AM is still significant,but this time, YEAR is also significant. This is due to the fact that we were testing YEAR as continuous previously.

#Question 9
Mr. Cheapbone should only sell tickets on days where visibility is good at 8am (and potentially run his operation only on "good" years).
