---
title: "Practical II"
author: "Katherine Filpo"
date: "5/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
apple <- read_csv("apple.csv")
#View(apple)
```

### Word Equation

Apple Rust ~ Wrapping + Temperature +Wrapping:Temperature
 OR
Virus ~ Wrapped + Temp + Wrapped:Temp

### Hypothesis 

*H~01~:* The interaction between wrapping and temeperature of the apple will not affect the apple rust. 

*H~02~:* The temperature of the apple will not affect apple rust.

*H~03~:* The wrapping of the apple will not affect apple rust.

### Full Analysis

Temp is catergorical variable:
```{r}
model_1 <-lm(VIRUS~TEMP+WRAPPED+WRAPPED:factor(TEMP),data=apple)
anova(model_1)
```

Temp is a continuous variable:
```{r}
model_2<- lm(VIRUS ~ factor(WRAPPED) * TEMP, data = apple)
anova(model_2)
```

No interaction betwen temp and wrapping:
```{r}
model_3<- lm(VIRUS ~ factor(WRAPPED) + factor(TEMP), data = apple)
anova(model_3)
```


***Model Opimizing***
```{r}
anova(model_1, model_2)
```

```{r}
anova(model_1, model_3)
```

**Model 1** is the better model because the Residual Sum of Squares is smaller. 

***Testing Residuals and Homogeneity of Variance***
```{r}
unstandardizedPredicted <- predict(model_1)
unstandardizedResiduals <- resid(model_1)

standardizedPredicted <- (unstandardizedPredicted - mean(unstandardizedPredicted)) / sd(unstandardizedPredicted)
standardizedResiduals <- (unstandardizedResiduals - mean(unstandardizedResiduals)) / sd(unstandardizedResiduals)

hist(standardizedResiduals, freq = FALSE)
curve(dnorm, add = TRUE) 
```

```{r}
plot(model_1, which =c(1,1))
```

The assumptions of Model_1 are sound. The standardized residuals and homogenitity of variance are as expected. 


### Conclusions + Graphs

```{r}
Fig1<-ggplot(apple, aes(x=TEMP, y=VIRUS,colour=WRAPPED, group=WRAPPED)) +
  geom_boxplot()+
  theme_classic()+
  xlab("Temperature")+ylab("Rot") +
  theme_classic() 
Fig1
```
 
 **Figure 1.** Apple Rot by Temperature and Wrapping. The affect of temperature on apple rot is significant (**P** = 0.039).*H~01~:* must be rejected, meaning that the temperature and wrapping of the apple is significant in terms of how much rust the apple will have. 

I would recommend that the Apple Orchardist American Association store their apples at a cooler temperature. Whether they are wrapped or not does not have a significant effect on the apple rot levels. 

