---
title: "Bio244 - LAB 9 - Two-Way Anova & Interactions"
author: "Gabriel Perron"
date: "2/25/2019"
output:
  #pdf_document: default
  html_document: default
---

```{r loading necessary datasets and packages, echo=FALSE, results="hide"}
#Before starting, I import my data
#RADISH <- read.csv("~/Dropbox/1_University/2_Education/1_Teaching/1_Bard/1_Courses/2_Biostats/0_Data_set/RADISH.csv")
RADISH <- read.csv("~/Documents/BioStats/lab9/RADISH.csv")
library(ggplot2)
library(gridExtra)
```


# Objective
Learning how to use two-way anova to test for multiple explanatory variables,

## Question 1
H~0i~: There is no effect of the interaction betwene soil type and water treatment on shoot length;  
H~Ai~: There is an effect of the interaction betwene soil type and water treatment on shoot length;   
H~01~: There is no effect of soil type on shoot length;  
H~A1~: There is an effect of soil type on shoot length;  
H~02~: There is no effect of water treatment on shoot length;  
H~A~: There is an effect of water treatment on shoot length;  
```{r Question 1}
model1 <- lm(RADISH$GREEN~factor(RADISH$SOIL)+factor(RADISH$WATER)+factor(RADISH$SOIL):factor(RADISH$WATER))
anova(model1)
```
**Conclusion**: There is a significant effect of the interaction betwene soil type and water treatment on shoot length (*F*~(1,60)~ = 19.359; *P* < 0.0001).  

```{r Figure 1}
(Figure1 <- ggplot(RADISH, aes(y=GREEN, x=factor(WATER), fill=SOIL)) +
            geom_boxplot() +
            xlab("Water Treatment") +
            ylab("Shoot lenght (cm)")+
            theme_classic())

#Now let's order water treatment from low to high on the x-axis
(Figure1_ordered <- ggplot(RADISH, aes(y=GREEN, x=factor(WATER, level = c('low','High')), fill=SOIL)) +
            geom_boxplot() +
            xlab("Water Treatment") +
            ylab("Shoot lenght (cm)")+
            theme_classic())

```
  
**Figure 1.** There is a significant effect of the interaction betwene soil type and water treatment on shoot length (*F*~(1,60)~ = 19.359; *P* < 0.0001). 


## Question 2
H0~i~: There is no effect of the interaction betwene soil type and water treatment on root length;  
H~A~: There is an effect of the interaction betwene soil type and water treatment on root length;   
H~01~: There is no effect of soil type on root length;  
H~A1~: There is an effect of soil type on root length;  
H~02~: There is no effect of water treatment on root length;  
H~A2~: There is an effect of water treatment on root length;  
```{r Question 2}
model1 <- aov(RADISH$ROOT~factor(RADISH$SOIL)+factor(RADISH$WATER)+factor(RADISH$SOIL):factor(RADISH$WATER))
summary(model1)
```
**Conclusion**: There is no effect of the interaction betwene soil type and water treatment on shoot length (*F*~(1,60)~ = 1.929; *P* = 0.17). However, there is an effect of soil type (*F*~(1,60)~ = 11.859; *P* = 0.001) and water treatment (*F*~(1,60)~ = 31.82; *P* < 0.0001) on shoot length.


```{r Figure 2}
Figure2A <- ggplot(RADISH, aes(x=factor(WATER, level =c('low','High')), y=ROOTS)) +
            geom_boxplot() +
            xlab("Water Treatment") +
            ylab("Root lenght (cm)")
Figure2B <- ggplot(RADISH, aes(x=SOIL, y=ROOTS)) +
            geom_boxplot() +
            xlab("Soil Type") +
            ylab("Root lenght (cm)")

grid.arrange(Figure2A, Figure2B, ncol=2)

```

**Figure 2.** There is no significant effect of the interaction betwene soil type and water treatment on shoot length (*F*~(1,60)~ = 19.359; *P* < 0.0001). However, there is an effect of **A)** water treatment (*F*~(1,60)~ = 31.82; *P* < 0.0001) and **B)** soil type (*F*~(1,60)~ = 11.859; *P* = 0.001) on shoot length.

## Question 3
To maximize growht of leaves and roots system, I would recommend using an organic soil with a high water treatment.



