---
title: "Lab 9 - Analysis of Variance"
author: "Katherine Filpo"
date: "3/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
RADISH <- read.csv("~/Documents/BioStats/lab9/RADISH.csv")
#to look at colors: colors()
```

## Lab 9: Interactions

### Question 1 - effects on shoot length

H~0~: There is no effect of water treatment or soil on shoot length.  

H~A~: There is an effect of water treament or soil on shoot length. 

**was supposed to have 6 total hypothesis here and for question 2
green~soil+water+soil:water
```{r}
model_1 <-aov(RADISH$GREEN ~ factor(RADISH$WATER)+factor(RADISH$SOIL) +factor(RADISH$SOIL:RADISH$WATER))
summary(model_1)
```

```{r pressure, echo=FALSE}
Fig1 <-ggplot(RADISH, aes(y = GREEN, x=factor(WATER, level = c('low', 'High')), fill = SOIL))+
  geom_boxplot() + 
  xlab("Water Treatment") + 
  ylab("Shoot Length (cm)") +
  theme_classic()

Fig1
#the level thing is putting the x axis in the correct order 
```

**Figure 1**.Water treatment and soil type has an effect on shoot length (F(1,60)=19.359; P < 0.0001), when taking into account the soil.
(the interaction is most important first, and since it is sigficicant, the other hypothesis are discarded)

**Conclusion** After accounting for the soil, there is an effect of water treament on shoot length, so we reject the H~0~ and say that the H~A~ is supported (*F*~(1,60)~=19.359; *P* < 0.0001). 

### Question 2 - effects on root

H~0~: There is no effect of water treatment on root length.

H~A~: There is an effect of water treatment on root length.

roots~soil+water+soil:water
```{r}
model_2 <-aov(RADISH$ROOTS ~ factor(RADISH$WATER)+factor(RADISH$SOIL) +factor(RADISH$SOIL:RADISH$WATER))
summary(model_2)
```

```{r, echo=FALSE}
Fig2 <- ggplot(RADISH, aes(y = ROOTS, x=factor(WATER, level = c('low', 'High')), fill = SOIL))+
  geom_boxplot() + 
  xlab("Water Treatment") + 
  ylab("Root Length (cm)") +
  theme_classic()
Fig2
```

**Figure 2**. Water treatment does not have an effect on root length (F(1,60)=1.929; P = 0.1700), when taking into account the soil.

**Conclusion** There is no sigficifant effect on the root length by the water treamentment when taking the soil into account, so we accept the H~0~. 

### Question 3
Taking into account all the statistical analysis performed, I would say that the best treatment for the maximum legnt of shoots and roots of the plant is the higher water treatment. Soil does play a slight role in this but can overall be considered to be a lesser effect than the water treatment. 

