---
title: "Lab3-Z_Test-Katherine_Filpo"
author: "Katherine Filpo"
date: "2/5/2020"
output: html_document
---

```{r setup, include=FALSE}
#library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
HERRING <- read.csv("~/Documents/BioStats/lab3/HERRING.csv") 
```

## Q1: Hypothesis 
Null hypothesis: there is no change of the body length and body depth in the fish of tivoli bay and the north atlantic population.

Hypothsis: there is a differemce between the north atlantic herrin and the tivoli bay herring 

Part a: mean body length
(of tivoli fish)
```{r cars}
mean(HERRING$LENGTH)
```

Part b: mean body depth
```{r}
mean(HERRING$DEPTH)
```
## Q2: Creating Z test
```{r}
z.test = function(a, mu, SD){
  z.statistic = (mean(a) - mu)/ SD
  return(z.statistic)
}
```

## Q3: Using the z test

Part a, testing for length
```{r}
z.test(mean(HERRING$LENGTH), 5.3, 2.7)
```

Part b, testing for depth
```{r}
z.test(mean(HERRING$DEPTH), 0.89, 0.11)
```
## Q4: Conclusions
Yes, I would say that I would reject the null hypothesis and that the hypothesis that there has been evolution is supported. The reason that I would say that is because the z test for the depth gives you the result of 5.485646 which is much higher than 1.96, meaning that it is out of the normal distribution, and therefore out of the norm. 

