---
title: "Bio244 - LAB 12 & 13 - Model Selection"
author: "Gabriel Perron"
date: "2019/4/01"
output:
  html_document: default
  pdf_document: default
---


```{r loading necessary datasets and packages, echo=FALSE, results="hide"}
#Before starting, I import my data
potatoes <- read.csv("~/Dropbox/1_University/2_Education/1_Teaching/1_Bard/1_Courses/2_Biostats/0_Data_set/potatoes.csv")
library(ggplot2)
library(dplyr)
```

# Question 1
Given you level of expertise at this point, I would say it is okay to only write down the null hypotheses when you have several pairs of hypotheses to write. I think by now, we all know that the alternative is simply the opposite.

**Hypotheses:**  
**H~01~**: There is no effect of the interaction between BAC, TEMP & OXYGEN on Rotting.  
**H~02~**: There is no effect of the interaction between BAC & TEMP on Rotting.  
**H~03~**: There is no effect of the interaction between BAC & OXYGEN on Rotting.  
**H~04~**: There is no effect of the interaction between OXYGEN & TEMP on Rotting.  
**H~05~**: There is no effect of BAC on Rotting.  
**H~06~**: There is no effect of TEMP on Rotting.  
**H~07~**: There is no effect of OXYGEN on Rotting.  
  
**Statistical analysis**
Here, I use the "*" as a shortcut to test the full statistical model. The stars basically include all the terms and all their interactions.
```{r}
model_1 <- lm(ROT~factor(BAC)*factor(TEMP)*factor(OXYGEN), data=potatoes)
anova(model_1)

#Alternatively you could write:
#model_1<- lm(ROT~factor(BAC)+factor(TEMP)+factor(OXYGEN)+factor(BAC):factor(TEMP)+factor(BAC):factor(OXYGEN)+factor(TEMP):factor(OXYGEN)+factor(BAC):factor(TEMP):factor(OXYGEN))
```
  
**Conclusions:**   
**H~01~**: We fail to reject. There is no effect of the interaction between BAC, TEMP & OXYGEN on Rotting (*F*~(4,36)~ = 0.869; *P* = 0.49).  
**H~02~**: We reject. There is an effect of the interaction between BAC and TEMP on Rotting (*F*~(2,36)~ = 3.27; *P* = 0.0498).  
**H~03~**: We fail to reject. There is no effect of the interaction between BAC & OXYGEN on Rotting (*F*~(4,36)~ = 0.321; *P* = 0.862).  
**H~04~**: We fail to reject.There is no effect of the interaction between OXYGEN & TEMP on Rotting (*F*~(2,36)~ = 0.34; *P* = 0.967).  
**H~05~**: We reject. There is an effect of BAC on Rotting (*F*~(2,36)~ = 13.912; *P* < 0.0001).  
**H~06~**: We reject. There is an effect of TEMP on Rotting (*F*~(2,36)~ = 36.202; *P* < 0.0001).  
**H~07~**: We fail to reject. There is no effect of OXYGEN on Rotting (*F*~(2,36)~ = 2.088; *P* = 0.14).  
  
**Data Visualization**
Here, because we only have a single significant interaction, it makes sense to plot the effect of BAC and TEMP. I decided to put BAC on the x-axis for reasons that should become obious in PART II. If you put TEMP on the x-axis for now, it is also good.
```{r}
ggplot(potatoes, aes(x=factor(BAC),y=ROT,fill=factor(TEMP))) +
            geom_boxplot() +
            xlab("Level of Bacteria") +
            ylab("Rotting")+
            theme_classic()

```
  
  **Figure 1.** There is an effect of the interaction between BAC and TEMP on Rotting (F~(2,36)~ = 3.27; *P* = 0.0498).
  
# Question 2

**normality of errors**
```{r}
#get unstandardized predicted and residual values
unstandardizedPredicted <- predict(model_1)
unstandardizedResiduals <- resid(model_1)

#get standardized values
standardizedPredicted <- (unstandardizedPredicted - mean(unstandardizedPredicted)) / sd(unstandardizedPredicted)
standardizedResiduals <- (unstandardizedResiduals - mean(unstandardizedResiduals)) / sd(unstandardizedResiduals)

#create residuals histogram
hist(standardizedResiduals, freq = FALSE)
curve(dnorm, add = TRUE) #run the following line at the same time as the one above.
```
The errors are pretty well distributed. Not biases on one side or the other. Let's look at the homogeneity of variance before making a final call.

**Homogeneity of errors**
```{r}
plot(model_1, which=c(1,1))
```
The residuals are evenly distributed around the center and do not follow any trend (e.g. redl ine is straight). Given the nice spread of the homogeneity of variance, I am happy with the normality of errors as well. Let's move on!

# Question 3 & Question 4

We first simplify the model by removing the three-way interaction. Remember, we can only remover terms from one level of complexity at the time.
```{r}
model_2 <- lm(ROT~factor(BAC)+factor(TEMP)+factor(OXYGEN)+factor(BAC):factor(TEMP)+factor(BAC):factor(OXYGEN)+factor(TEMP):factor(OXYGEN), data=potatoes)
anova(model_2)
```
Here, I will stop stating the conclusions. The truth is that when one is busy working on model simplification, it is easier to simply modify as much as possible until you choose a model to state you final conclusions. The idea is that you don't want to loose your train of thougt while making the different decisions.

Now,  let's compare the two models to each other using the anova() function as shown in the lecture, testing the following two hypotheses:
**H~0~**: There is not difference in the variability explained between the two models.  
**H~A~**: There is a difference in the variability explained between the two models.  
```{r}
anova(model_1,model_2)
```
We reject **H~0~**: The The RSS is not different between model_1 and model_2, therefore we pick model_2 - the simpler one.  
  
We will then further simplify the model by removing the two-way interactions that are not significant:
```{r}
model_3 <- lm(ROT~factor(BAC)+factor(TEMP)+factor(OXYGEN)+factor(BAC):factor(TEMP), data=potatoes)
anova(model_3)
```
  
Once again, we will compare the two models for model fit:  
**H~0~**: There is not difference in the variability explained between the two models.  
**H~A~**: There is a difference in the variability explained between the two models.  
```{r}
anova(model_2, model_3)
```
We reject **H~0~**: The The RSS is not different between model_2 and model_3, therefore we pick model_3 - the simpler one.  
  
We further simplify the model by remocing OXYGEN from the model altogether:
```{r}
model_4 <- lm(ROT~factor(BAC)*factor(TEMP), data=potatoes)
anova(model_4)
```
  
We then compare the models for model fit:  
**H~0~**: There is not difference in the variability explained between the two models.  
**H~A~**: There is a difference in the variability explained between the two models.  
```{r}
anova(model_3, model_4)
```
We reject **H~0~**: The RSS is not different between model_3 and model_4, therefore we pick model_4 - the simpler one.


Finally, since the two-way interaction is not "super" significant, we might be tempted to remove it as well. Let's give it a try:
```{r}
model_x <- lm(ROT~factor(BAC)+factor(TEMP), data=potatoes)
anova(model_x)
```
Of course, BAC and TEMP are still very significant, but let's see if removing the significant interaction improved on the model:
**H~0~**: There is not difference in the variability explained between the two models.  
**H~A~**: There is a difference in the variability explained between the two models.  
```{r}
anova(model_4,model_x)
```
Ha ha! here we see that we reject H~0~; there is a difference in the variability explaiend between the two models. Because Model 1 (or model 4) has a lower RSS, this means that it explains significantly more variability and we should keep it.

              
# Question 5
Considering BAC as continuous could be a good idea because BAC seems to follow a linear trend
when looking at Figure 1.
              

# Question 6
We will start by removing the three-way interactions:
```{r}
model_5 <- lm(ROT~BAC*factor(TEMP)*factor(OXYGEN), data=potatoes)
anova(model_5)
```



# Question 7 & 8
```{r}
model_6 <- lm(ROT~BAC+factor(TEMP)+factor(OXYGEN)+BAC:factor(TEMP)+BAC:factor(OXYGEN)+factor(TEMP):factor(OXYGEN), data=potatoes)
anova(model_6)
```

Once again, we will compare the two models for model fit: 
```{r}
anova(model_5,model_6)  
```
The RSS is not different between model_5 and model_6, therefore we pick model_6 - the simpler one.
  
We then removed the two-way intereactions that are not significant:
```{r}
model_7 <- lm(ROT~BAC+factor(TEMP)+factor(OXYGEN)+BAC:factor(TEMP), data=potatoes)
anova(model_7)
```

    
Once again, we will compare the two models for model fit: 
```{r}
anova(model_6,model_7)
```
The RSS is not different between model_6 and model_7, therefore we pick model_7 - the simpler one.
  
We remove Oxygen from the model altogher:
```{r}
model_8 <- lm(ROT~BAC+factor(TEMP)+BAC:factor(TEMP), data=potatoes)
anova(model_8)
```

  
Once again, we will compare the two models for model fit: 
```{r}
anova(model_7,model_8) 
```
The RSS is not different between model_7 and model_8, therefore we pick model_8 - the simpler one.
  
Finally, let's check whether it is best to consider BAC as continuous or categorical:
```{r}
anova(model_4,model_8)
```
The RSS is not different between model_4 and model_8. We pick model_8 which has more DF in residual. It is thus a good idea to consider BAC as continuous in this model.

              
# Question 9
```{r}
ggplot(potatoes, aes(x=BAC, y=ROT,colour=factor(TEMP), group=TEMP)) +
  geom_point()+
  geom_smooth(method=lm,se=FALSE)+
  theme_classic()+
  xlab("Bacteria Level")+ylab("Number of rotten potatoes")+
  xlim("Low", "Medium", "High") +
  theme_classic() +
  theme(axis.line.x = element_line(color = "black"), axis.line.y = element_line(color = "black"))
```
              
**Figure 2**. The effect of bacterial load on potatoes spoilage at different temperature. We found that the initial bacterial load significant predicted spoilage of different temperature (*F*~(1,50)~ = 6.98; *P* < 0.01). 


# Question 10
I would recommend to store potatoes in a cold room and to minimize the amount of bacteria present on the potatoes at the time of storage.

# Question 11

Becfore you can use filter(), you have to install the dyplr package (see in the first chunk of this markdown).
```{r}
temp2 <- filter(potatoes, TEMP =="2")
temp2_lm <- lm(data=temp2, ROT~BAC)
anova(temp2_lm)
```

  
```{r}
summary(temp2_lm)
```
**Equation:** Y = 6.23X + 0.81; X^2^ = 0.50


```{r}
ggplot(temp2, aes(x=BAC, y=ROT)) +
  geom_point()+
  geom_smooth(method=lm,se=FALSE)+
  theme_classic()+
  xlab("Bacteria Level")+ylab("Number of rotten potatoes")+
  xlim("Low", "Medium", "High") +
  theme_classic() +
  theme(axis.line.x = element_line(color = "black"), axis.line.y = element_line(color = "black"))
```
  
**Figure 3.** The effect of bacterial load on potatoes spoilage at temperature 2 (*F*~(1,25)~ = 26.67; *P* < 0.0001). The regression line is explained by Y = 6.23X + 0.81; X^2^ = 0.50.



Our model indeed predicts that spoilage would further diminish as the initial bacterial load decreases. I would suggest taking great care of reducing the initial bacterial load on the potatoes. Maybe washing the legumes before storing.




