---
title: "lab11 - Transforming Data"
author: "Katherine Filpo"
date: "4/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(readxl)
bristles <- read_excel("bristles.xlsx")
```

### Question 1

I expect that the data requires a transformation because the experiment will most likely cause a very drastic change in the flies, as mutations can be small or very large.

### Question 2  - Untransformed data set

H~0~ The different levels of radation will not have an effect on the brisltes of the flies.

H~A~ The different levels of radation will have an effect on the bristles.

```{r, echo = FALSE}
model_1 <- aov(bristles$NBRIST~factor(bristles$GEN)+factor(bristles$BLOCK))
summary(model_1)
```

**Conclusion**
The H~A~ is supported (**F**(4,20)=36.24). 

### Question 3
The homogeneity of variance seems to be well distributed as the data set makes a normal distribution shape. 

```{r}
unstandarizedPred <- predict(model_1)
unstandarizedRes <- resid(model_1)

standarizedPred <- (unstandarizedPred - mean(unstandarizedPred)) / sd(unstandarizedPred)
standarizedRes <- (unstandarizedRes - mean(unstandarizedRes)) / sd(unstandarizedRes)

hist(standarizedRes, freq = FALSE)
curve(dnorm, add = TRUE)
```

```{r}
par(mfrow=c(2,2))
plot(model_1)
```

### Question 4  - Transformed data set: Square Root
```{r, echo=FALSE}
model_2 <- aov(bristles$NBRISTsqrt~factor(bristles$GEN)+factor(bristles$BLOCK))
summary(model_2)
```

**Conclusion**
The H~A~ is supported (**F**(4,20)=57.38). The distribution here also seem to be well as there is a normal distribution, and a flatter curve. 
```{r}
unstandarizedPred <- predict(model_2)
unstandarizedRes <- resid(model_2)

standarizedPred <- (unstandarizedPred - mean(unstandarizedPred)) / sd(unstandarizedPred)
standarizedRes <- (unstandarizedRes - mean(unstandarizedRes)) / sd(unstandarizedRes)

hist(standarizedRes, freq = FALSE)
curve(dnorm, add = TRUE)
```


### Question 5  - Transformed data set: Natural Log
```{r, echo=FALSE}
model_3 <- aov(bristles$NBRISTln~factor(bristles$GEN)+factor(bristles$BLOCK))
summary(model_3)
```

**Conclusion**
The H~A~ is supported (**F**(4,20)=27.12). The distribution here also seem to be well as there is a normal distribution, and one can see the flattest curve out of all three in this analysis. 
```{r}
unstandarizedPred <- predict(model_3)
unstandarizedRes <- resid(model_3)

standarizedPred <- (unstandarizedPred - mean(unstandarizedPred)) / sd(unstandarizedPred)
standarizedRes <- (unstandarizedRes - mean(unstandarizedRes)) / sd(unstandarizedRes)

hist(standarizedRes, freq = FALSE)
curve(dnorm, add = TRUE)
```

### Final Figure
```{r}
ggplot(bristles, aes(x=factor(GEN), y=bristles$NBRISTln)) +
            geom_boxplot() +
            xlab("UV Radation Level") +
            ylab("Bristle per Fly") +
            theme_classic()  
```

**Figure 1**: The **F**(4,20) is 27.12, so it is statistically significant. 

