---
title: "Lab7_Variance_Analysis"
author: "Katherine Filpo"
date: "2/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
HUMAN <- read.csv("~/Documents/BioStats/lab7/human.csv")
View(HUMAN)
library(ggplot2)
```
H~0~ The index finger length and the height of a person is completely unrelated. 

H~A~ Based on the scatter plots, the longer the index finger, the taller the height of the person, therefore, the index finger is the best predictor of the height. 

## Q1 PLot of Forearm and Index fingers

```{r pressure, echo=FALSE}
ggplot(data = HUMAN) +geom_point(mapping = aes(x=FORE, y=HEIGHT, color = 'red'))
```

```{r, echo=FALSE}
ggplot(data = HUMAN) +geom_point(mapping = aes(x=INDEX, y=HEIGHT)) 
##+ guide = guide_legend figure out how to add this
```

## Anova for Forearm predicting 
```{r}
model_1<-lm(HUMAN$HEIGHT ~ HUMAN$FORE)
anova(model_1)
summary(model_1)
```
```{r}
ggplot(HUMAN, aes(x = HEIGHT, y = FORE)) +geom_point() + geom_smooth(method=lm)
```

```{r}
model_2<-lm(HUMAN$HEIGHT ~ HUMAN$INDEX)
anova(model_2)
summary(model_2)
```

```{r}
ggplot(HUMAN, aes(x = HEIGHT, y = INDEX)) +geom_point() + geom_smooth(method=lm)
```

##Final Conclusions
From the anovas, it is most clear that the forearm length is a better measurement to be used in predicting height of humans, so we accept the null hypothesis. 
