---
title: "Bio244 - LAB 10 - Independence"
author: "Gabriel Perron"
date: "Updated on 2020/3/31"
output:
  html_document: default
  pdf_document: default
---


```{r, echo=FALSE, results="hide",message=FALSE,warning=FALSE}
#loading necessary libraries for running the script
library(ggplot2)
library(gridExtra)
library(grid)
library(dplyr)
```

# Question 1
Dr. Sharp would  conclude that there is a significant difference in the lookup rate betwen make and female sheep.


# Question 2 & 3
Here, the individual data points are not independent from each other. Dr. Sharp took multiple (20) measurement from each sheep. In reality, there is only 6 sheep in the experiment, or 6 independent data point. I would suggest taking the average lookup rate for each individual sheep. To do so, I would either compute the average for each sheep and create a new csv file with the six datpoint. Or it is possible to use R to create a new table using the following code:



```{r}
sheep <- read_excel("sheep.xlsx")

sheepavg <- sheep %>%  #Create a new dataframe named sheepavg
  group_by(SHEEP) %>%  #Group the values by sheep
  mutate(avgNLOOKUPS = mean(NLOOKUPS)) %>% #Create a new column for averages
  distinct(avgNLOOKUPS,SHEEP,SEX) #Keep one value per sheep

g=tableGrob(sheepavg,rows = NULL)
grid.newpage()
grid.draw(g)


```
 
 
 **Table 1.** Average lookup rate for each sheep

*N.B. I do not expext you to be able to do so in a practical exam. But I want to show you the code in case you would like to learn how to use R for more functions.*


# Question 4

**a) Experimental Design:**
```{r}
Table2 = data.frame(Male=3,
                   Female=3)
g=tableGrob(Table2, rows = NULL)
grid.newpage()
grid.draw(g)
```

  **Table 2.** Experimental Diagram

**b) Word equation:** AVGLOOKUPRATE ~ SEX

**c) Hypotheses:**

**H~0~**: There is no different in average loopup rate between male and female sheep.
**H~A~**: There is a different in average loopup rate between male and female sheep.

**d) Statisitcal Analysis:**
```{r Figure 1}
model_sheep <- lm(sheepavg$avgNLOOKUPS~factor(sheepavg$SEX))
anova(model_sheep)
```

**e) Conclusion:** 
We fail to reject H~0~. There is no different in average loopup rate between male and female sheep (*F*~(1,4)~ = 1.802; *P* = 0.25).


**f) Data Vizualisation:**
```{r}
ggplot(sheepavg, aes(x=factor(SEX), y=avgNLOOKUPS)) +
            geom_boxplot() +
            xlab("Sheep sex") +
            ylab("Average Lookup Rate") +
            theme_classic() 
```
 
 **Figure 1.** There is no different in average loopup rate between male and female sheep (*F*~(1,4)~ = 1.802; *P* = 0.25).

# Question 5
I would suggest that Dr. Sharp re-allocate his time studying more sheep rather than studying the same sheep over and over.






